Shapes analysis for time series.

### 2024-05-21 NEURIPS24/paper.tex
update

 %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
 %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
 %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
 %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
 %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
 Balancing between discrete and continuous elements is a challenging task.
  In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analogue to Residual Networks \cite{zagoruyko2016wide}.
   This approach was further enhanced by Neural Controlled Differential Equations (Neural CDEs) \cite{kidger2020neural} for handling irregular time series, functioning as continuous-time analogs of RNNs \cite{schuster1997bidirectional}.
 Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
  Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
  Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
  With TS-LDDMM, the representation is also driven by an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm.

>>>>

 %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
 %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
 %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
 %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
 %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
 Balancing between discrete and continuous elements is a challenging task.
  In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analogue to Residual Networks \cite{zagoruyko2016wide}.
   This approach was further enhanced by Neural Controlled Differential Equations (Neural CDEs) \cite{kidger2020neural} for handling irregular time series, functioning as continuous-time analogs of RNNs \cite{schuster1997bidirectional}.
 Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
  Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
  Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
  With TS-LDDMM, the representation is also derived from an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm, which enhance interpretability.
